---
title: "Getting Started with reedling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with reedling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

This vignette walks through the core reedling workflow: loading ReEDS run data,
formatting technology categories, and creating common plots.

## Setup

```{r libraries, message = FALSE}
library(reedling)
library(data.table)
library(ggplot2)
library(ggthemes)
```

## Loading runs

Use `run_summary()` to scan a directory of ReEDS runs. Here we point to the
example runs bundled with the package:

```{r load-runs}
runspath <- system.file("extdata", "runs", package = "reedling")
runs <- run_summary(runspath, "20260810")
runs
```

## Installed capacity

Load capacity data from `outputs.h5` and apply standard technology formatting:

```{r cap}
cap <- load_h5_data(runs, "cap")
cap <- format_techs(cap)
```

Plot installed capacity by technology and run:

```{r cap-plot, fig.width = 8, fig.height = 5}
cap_agg <- cap[, by = .(run, t, tech), .(cap_MW = sum(cap_MW))]

ggplot(cap_agg[t >= 2025], aes(x = factor(t), y = cap_MW / 1e3, fill = tech)) +
  geom_col() +
  facet_wrap(~run, nrow = 1) +
  scale_x_discrete("") +
  scale_y_continuous("Installed capacity [GW]",
    expand = expansion(mult = c(0, 0.05))) +
  scale_fill_manual(values = tech_colors) +
  guides(fill = guide_legend(ncol = 1))
```

## Annual generation

Load annual generation data with a different technology aggregation:

```{r gen}
gen_ann <- load_h5_data(runs, "gen_ann")
gen_ann <- format_techs(gen_ann, col_to = "agg3")
```

## Transmission capacity

Load transmission data and plot total capacity over time:

```{r transmission}
tran_mi_out <- load_h5_data(runs, "tran_mi_out")
tran_mi_out$trtype <- factor(tran_mi_out$trtype, levels = c("LCC", "B2B", "AC"))
tran_mi_out_total <- tran_mi_out[, by = .(run, t),
  .(`tran_mi_out_MW-mi` = sum(`tran_mi_out_MW-mi`))]
```

```{r transmission-plot}
ggplot(tran_mi_out_total[t >= 2025],
  aes(x = t, y = `tran_mi_out_MW-mi` / 1e6, color = run)) +
  geom_line() +
  scale_y_continuous("Transmission capacity [TW-mi]",
    expand = expansion(mult = c(0, 0.05)), limits = c(0, NA)) +
  scale_x_continuous("", expand = expansion(mult = c(0, 0.1)))
```

## Bokeh report data

You can also load data from bokeh report Excel files. First, preview available
fields:

```{r bokeh-fields}
list_bokeh_fields(runs)
```

Then load a specific result, such as system cost:

```{r npv}
npv <- load_bokeh_data(runs, "29_Present Value of System Cost")
npv_net <- npv[, by = .(run), .(net = sum(`Discounted Cost (Bil $)`))]
```

```{r npv-plot, fig.width = 8, fig.height = 5}
ggplot(npv, aes(x = run)) +
  geom_col(mapping = aes(y = `Discounted Cost (Bil $)`, fill = cost_cat)) +
  geom_point(data = npv_net, mapping = aes(y = net)) +
  geom_label(data = npv_net, mapping = aes(y = net, label = round(net)),
    vjust = -0.5, size = 2.5) +
  geom_hline(yintercept = 0) +
  scale_x_discrete("") +
  scale_fill_tableau(palette = "Tableau 20")
```

## Loading CSV outputs

In addition to `outputs.h5`, you can load individual CSV files with
`load_run_data()`:

```{r csv}
cap_csv <- load_run_data(runs, "cap.csv")
hierarchy <- load_run_data(runs, "hierarchy.csv", folder = "inputs_case")
```
